---
title: "Choosing a data-set for downstream analysis"
output: html_document
---

After CEL-seq and DEG analysis, we got two datasets. The first one denoted some of the adjusted p-value as *NA*. The second dataset assigned a valid adjusted p-value to each measured gene, which we received upon specifically asking for this feature.

The purpose of this analysis is to estimate the validity of using the dataset with the assigned adjusted p-values.

First, let's read in the two datasets.
```{r}
N2.excl.low <- read_csv("raw/2018-04-25 reanalysis results Eitan Winter/diff_exp_WT_anc1/WT.EV_WT.anc1_results.csv",
                        col_types = cols(NCBI_gene_ID = col_character()))
N2.excl.non <- read_csv("raw/2018-05-29 reanalysis results Eitan Winter without excluding low intensity/diff_exp_WT_anc1/WT.EV_WT.anc1_results.csv",
                        col_types = cols(NCBI_gene_ID = col_character()))
```

We'll reorder the gene names so that the order of the two datasets match.
```{r}
N2.excl.low <- arrange(N2.excl.low,Gene_stable_ID)
N2.excl.non <- arrange(N2.excl.non,Gene_stable_ID)
```

We'll compare the gene IDs row-by-row in the two dataset to validate that their order isidentical (mean equaling 1 indicates that all comparisons are TRUE)
```{r}
mean(N2.excl.low$Gene_stable_ID == N2.excl.non$Gene_stable_ID)
```

We'll confirm all gene IDs and positions match by plotting all the  data except adjusted p-values, expecting a complete linear correlation, which is indicated by the red line.
```{r}
par(mfrow = c(3,2))
plot(N2.excl.low$baseMean,N2.excl.non$baseMean)
abline(0,1,col = "red")
plot(N2.excl.low$log2FoldChange,N2.excl.non$log2FoldChange)
abline(0,1,col = "red")
plot(N2.excl.low$lfcSE,N2.excl.non$lfcSE)
abline(0,1,col = "red")
plot(N2.excl.low$baseMean,N2.excl.non$baseMean)
abline(0,1,col = "red")
plot(N2.excl.low$stat,N2.excl.non$stat)
abline(0,1,col = "red")
plot(N2.excl.low$pvalue,N2.excl.non$pvalue)
abline(0,1,col = "red")
```

Let's change NA values to -0.1 to be able to visualize them, and validate that the other dataset with assigned adjusted p-values, have no NA values.
```{r}
N2.excl.low$padj[is.na(N2.excl.low$padj)] <- -0.1
mean(is.na(N2.excl.non$padj))
```

Scatter plotting the adjusted p-values of the two datasets.
```{r}
par(mfrow = c(1,2))
plot(N2.excl.low$padj,N2.excl.non$padj)
abline(0,1,col = "red")
```

Focusing on the 0 to 0.1 p-values range.
```{r}
plot(N2.excl.low$padj,N2.excl.non$padj,
     xlim = c(0,0.1),
     ylim = c(0,0.1))
abline(0,1,col = "red")
```
This reveals that the newly assigned p-values span from 0 to 1. In addition, we see that by assigning the adjusted p-values instead of NA values, several genes that were initially assigned an adjusted p-value, do not pass the /alpha = 0.1 threshold of statistical significance.

Most genes reached a significance level > 0.1, still let's focus on those "new candidates" whose p-value is less than 0.1, and had NA values in the 2018-04-25 analysis.
```{r}
new.cand <- N2.excl.non[N2.excl.non$padj < 0.1 & N2.excl.low$padj == -0.1,]
par(mfrow = c(1,1))
plot(new.cand$baseMean,new.cand$log2FoldChange)
new.cand$Gene_name
```
 
Visualize how does the read count of these genes compares to all the other measured genes.
```{r}
par(mfrow = c(1,2))
plot(log(N2.excl.low$baseMean),
     N2.excl.low$log2FoldChange,
     xlab = "log(Reads)",
     ylab = "log2(fold-change)")
points(log(new.cand$baseMean),new.cand$log2FoldChange,col = "red")

plot(log(N2.excl.low$baseMean),
     N2.excl.non$padj,
     xlab = "log(Reads)",
     ylab = "Adjusted p-value")
points(log(new.cand$baseMean),new.cand$padj,col = "red")
```
It's clear that the genes that were set as NA are defined as such due to their low read count. This is could be necessary to avoid false positives.

# Summary
I will carry on the analysis using the data from 2018-04-24