---
title: "Fetching attributes from BioMart"
author: "Amir Levine"
date: "201-11-27"
output: md_document
---

# Pre-loading
```{r Load Packages}
source("https://bioconductor.org/biocLite.R")
if (!require(biomaRt)) {BiocManager::install("biomaRt"); library(biomaRt)}
if (!require(org.Ce.eg.db)) {BiocManager::install("org.Ce.eg.db"); library(org.Ce.eg.db)}
```

Fetch BioMart object and browse annotations.

```{r Fetch BioMart object}
mart = useMart(biomart="ensembl", dataset="celegans_gene_ensembl")
```

```{r Browse annotations}
att <- listAttributes(mart)
att[att$page == "feature_page",]
```
GO terms associated attributes:
```
                               name                                description         page
...
25                            go_id                          GO term accession feature_page
26                        name_1006                               GO term name feature_page
27                  definition_1006                         GO term definition feature_page
28                  go_linkage_type                      GO term evidence code feature_page
29                   namespace_1003                                  GO domain feature_page
30             goslim_goa_accession                    GOSlim GOA Accession(s) feature_page
31           goslim_goa_description                     GOSlim GOA Description feature_page
...
```

KEGG, MetaCyc associated attributes
```
                               name                                description         page
...
36                      kegg_enzyme                 KEGG Pathway and Enzyme ID feature_page
38                          metacyc                                 MetaCyc ID feature_page
```


Fetch Gene-Ontology (GO) and GO-Slim annotations for *C. elegans*. 
The fetching process takes a while, so save the data for faster future run-times.
```{r Fetch GO annotations}
att <- c("ensembl_gene_id",
         "go_id",
         "name_1006",
         "definition_1006",
         "go_linkage_type",
         "namespace_1003")
go_annotation <- getBM(attributes = att, 
                       filters=list(biotype='protein_coding'),
                       mart=mart)

att <- c("ensembl_gene_id",
         "goslim_goa_accession",
         "name_1006",
         "goslim_goa_description",
         "go_linkage_type",
         "namespace_1003")
goslim_annotation <- getBM(attributes = att, 
                       filters=list(biotype='protein_coding'),
                       mart=mart)

save(go_annotation,goslim_annotation, file = "data/2018-11-27-GO-annotations.Rdata")
```

# Entrez IDs for KEGG enrichment analysis
Set the  
Obtain all the *ensembl_gene_id* and *entrezgene* for *C. elegans*.
Save for later usage by enrichment analysis
```{r}
entrez <- getBM(
  attributes = c("ensembl_gene_id","entrezgene"),
  mart = mart
)
save(entrez, 
  file = "experiments/2018-05-01-kegg-enrichment/data/2018-11-29-entrez-ids-for-kegg-enrichment.Rdata")
```