---
title: "Enrichment of T-box transcription factors binding motifs in the promoter regions of anc-1 RNAi regulated genes"
author: "Amir Levine"
date: "2019-02-04"
output: md_document
---

```{r packages}
if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(PWMEnrich)) {BiocManager::install("PWMEnrich"); library(PWMEnrich)}
```

Load packages and previously generated data of T-box transcriptions factors enrichment in each DEG set.
```{r load-data}
load("data/2018-11-27-geneExpression-pval-thresh-0.05.rdata",verbose = TRUE) # CEL-seq processed data of anc-1 RNAi, p<0.05
load("data/Celegans-promoters.rdata",verbose = TRUE) # Promoter regions +500 to -100 bp of TSS
load("experiments/2018-07-11-tbx-motifs/data/backgound-celegans-proteincoding-tbx.rdata",verbose = TRUE) # PWMEnrich background T-box TFs
```

Focus on the intersect of the two strains, with upregulated and downregulated genes analyzed separately. 

## Downregulated genes 
```{r regulated-genes}
goi <- DEG$INTERSECT.dn # CEL-seq DEGs, query genes
length(goi) # 308 downregulated genes in both strains (w/ p-value < 0.05)
mRNA.seqs.promoter <- mRNA.seqs.promoter %>% mutate(CELseq.hits = ensembl_gene_id %in% goi)
```

Motif enrichment analysis in gene-sets
```{r enrichment}
useBigMemoryPWMEnrich(TRUE) # using large amount of memory
res.dn = motifEnrichment(prs[mRNA.seqs.promoter$CELseq.hits], bg.custom)
useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
```

Generate table of transcription factors that are enriched in the promoters of anc-1 downregulated genes. Export the reports as tables
```{r export-table}
report.dn <- groupReport(res.dn,by.top.motifs=FALSE)
plot(report.dn)
report.dn %>% 
  as.data.frame() %>% 
  mutate("enriched.motif" = report.dn$p.value < 0.05) %>%
  write_csv(path = "experiments/2018-07-11-tbx-motifs/results/2019-02-07-Tbox-motifs-enrichment-overlapping-downregulated.csv")
```

```{r save-plot}
#' Export the reports 
pdf(file = "experiments/2018-07-11-tbx-motifs/results/2019-02-04-Tbox-motifs-enrichment-overlapping-downregulated.pdf",
    height = 1.5,
    width = 5)
plot(report.dn, fontsize=7, id.fontsize=6)
dev.off()
```

## Upregulated genes 
```{r regulated-genes}
goi <- DEG$INTERSECT.up # CEL-seq DEGs, query genes
length(goi) # 338 downregulated genes in both strains (w/ p-value < 0.05)
mRNA.seqs.promoter <- mRNA.seqs.promoter %>% mutate(CELseq.hits = ensembl_gene_id %in% goi)
```

Motif enrichment analysis in gene-sets
```{r enrichment}
useBigMemoryPWMEnrich(TRUE) # using large amount of memory
res.up = motifEnrichment(prs[mRNA.seqs.promoter$CELseq.hits], bg.custom)
useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
```

Generate table of transcription factors that are enriched in the promoters of anc-1 downregulated genes. Export the reports as tables
```{r export-table}
report.up <- groupReport(res.up,by.top.motifs=FALSE)
plot(report.up)
report.up %>% 
  as.data.frame() %>% 
  mutate("enriched.motif" = report.dn$p.value < 0.05) %>%
  write_csv(path = "experiments/2018-07-11-tbx-motifs/results/2019-02-07-Tbox-motifs-enrichment-overlapping-upregulated.csv")
```

```{r save-plot}
#' Export the reports 
pdf(file = "experiments/2018-07-11-tbx-motifs/results/2019-02-04-Tbox-motifs-enrichment-overlapping-upregulated.pdf",
    height = 1.5,
    width = 5)
plot(report.up, fontsize=7, id.fontsize=6)
dev.off()
```

```{r save-data}
save(report.up,report.dn,res.dn, res.up,
     file = "experiments/2018-07-11-tbx-motifs/data/2019-02-07-Tbox-anc1-reg-enriched.Rdata")
```

