---
title: "Descriptive statistics and correlations of T-box transcription factors motifs in *anc-1* regulated genes"
author: "Amir Levine"
date: '2018-10-28'
output: md_document
---
Load packages and datasets.
PWMEnrich object of Tbox transcrption factors motifs enrichement in the promoters of *anc-1* RNAi regulated genes.
```{r}
if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(Hmisc)) {install.packages("Hmisc"); library(Hmisc)} # To calculate correlations and their p-values
if (!require(corrplot)) {install.packages("corrplot"); library(corrplot)} # To plot colored correlations.
if (!require(psych)) {install.packages("psych"); library(psych)}  # For correlation Matrix
library(RColorBrewer) # To plot colored correlations.

load("data/geneExpression.rdata")
load("experiments/2018-07-11-tbx-motifs/data/Tbox-anc1-reg-enriched.Rdata")
```

## Create a matrix of T-box TFs enrichment *p*-values for each gene regulated by *anc-1* RNAi
Obtain the *p*-values of the T-box PWM motifs per gene promoter sequence.  Rows correspond to the different input promoter sequences and the columns correspond to the motifs.  `sequence.bg` contains the corresponding *p*-values. (`sequence.nobg` contains the raw affinity scores).
```{r}
res <- list(res.dn,res.up,res.reg)
res.pvals <- sapply(res, function(x) x$sequence.bg)
names(res.pvals) <- c("res.dn","res.up","res.reg")
```

Rename the columns as T-box protein names.
```{r}
res.pvals <- lapply(res.pvals, function(x) {colnames(x) <- c("TBX-33","TBX-38","TBX-39","TBX-40","TBX-43"); x})
```

## Descriptive statistics of the T-box motifs in the *anc-1* regulated genes
Change *p*-value to Boolean. TRUE is statistically significant (*p*-value < 0.05).
```{r}
res.sig <- lapply(res.pvals,function(x) x < 0.05)
```

In what percentage of the sequences does each T-bpx transcription factor appear?
```{r}
mapply(function(p,sig) {round( colSums(sig) / nrow(p)  * 100, 2)},
       res.pvals, res.sig)
```
```
# % of genes for each T-box
       res.dn res.up res.reg
TBX-33   7.46   4.79    6.07
TBX-38   8.96   5.71    7.26
TBX-39   7.71   4.79    6.19
TBX-40   8.71   5.02    6.79
TBX-43   5.47   5.02    5.24
```
All T-box transcription factors appear to be present in 5%-9% of the *anc-1* regulated genes, more so in down-regulated genes.


## Correlations of *p*-values in all *anc-1* regulated genes
Is there a correlation in motif presence between any two motifs?
```{r}
res.cor <- lapply(res.pvals, Hmisc::rcorr)
print(res.cor)
#pvals.cor <- Hmisc::rcorr(pvals.dn.mat)
```
```
$`res.dn`
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.26   0.12   0.12   0.27
TBX-38   0.26   1.00   0.02   0.08   0.83
TBX-39   0.12   0.02   1.00   0.63   0.03
TBX-40   0.12   0.08   0.63   1.00   0.08
TBX-43   0.27   0.83   0.03   0.08   1.00

n= 402 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.0154 0.0133 0.0000
TBX-38 0.0000        0.6263 0.1289 0.0000
TBX-39 0.0154 0.6263        0.0000 0.4903
TBX-40 0.0133 0.1289 0.0000        0.1020
TBX-43 0.0000 0.0000 0.4903 0.1020       

$res.up
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.22   0.02   0.00   0.22
TBX-38   0.22   1.00   0.02   0.00   0.82
TBX-39   0.02   0.02   1.00   0.51   0.03
TBX-40   0.00   0.00   0.51   1.00   0.07
TBX-43   0.22   0.82   0.03   0.07   1.00

n= 438 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.7065 0.9966 0.0000
TBX-38 0.0000        0.6427 0.9335 0.0000
TBX-39 0.7065 0.6427        0.0000 0.4724
TBX-40 0.9966 0.9335 0.0000        0.1302
TBX-43 0.0000 0.0000 0.4724 0.1302       

$res.reg
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.25   0.07   0.06   0.25
TBX-38   0.25   1.00   0.03   0.05   0.82
TBX-39   0.07   0.03   1.00   0.57   0.04
TBX-40   0.06   0.05   0.57   1.00   0.08
TBX-43   0.25   0.82   0.04   0.08   1.00

n= 840 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.0391 0.0604 0.0000
TBX-38 0.0000        0.4365 0.1826 0.0000
TBX-39 0.0391 0.4365        0.0000 0.2963
TBX-40 0.0604 0.1826 0.0000        0.0208
TBX-43 0.0000 0.0000 0.2963 0.0208       

```

Plot correlations of *p*-values matrix.
```{r}
f_corrplot <- function(cor) {
  corrplot::corrplot(cor$r,
                     method = "shade",
                     type="upper",
                     addgrid.col = "black",
                     #order="hclust",
                     col=brewer.pal(n=10, name="RdBu"),
                     tl.col="black", 
                     tl.srt=45,
                     cl.align.text="l",
                     p.mat = cor$P,
                     sig.level = 0.05,
                     insig = "label_sig",
                     # hide correlation coefficient on the principal diagonal
                     diag=FALSE )
}

res.cor.plots <- lapply(names(res.cor), function(x) {
  svg(filename = paste0("experiments/2018-10-28-tbx-enrich-corr/results/tbx-motifs-cor-",x,".svg"),
      height = 2, width = 2, pointsize = 7)
  f_corrplot(res.cor[[x]])
  dev.off()
})
      

# # All down-regulated genes
# cp <- corrplot::corrplot(pvals.cor$r,
#                          method = "shade",
#                          type="upper",
#                          addgrid.col = "black",
#                          order="hclust",
#                          col=brewer.pal(n=10, name="RdBu"),
#                          tl.col="black", 
#                          tl.srt=45,
#                          cl.align.text="l",
#                          p.mat = pvals.cor$P,
#                          sig.level = 0.05,
#                          insig = "label_sig",
#                          # hide correlation coefficient on the principal diagonal
#                          diag=FALSE)
# dev.off()
```

## Descriptive statistics of the T-box motifs in the downregulated genes with at least one T-box motif

Filter-out genes w/o a motif.
```{r}
res.pvals.wmotif <- mapply(function(p,sig) subset(p,rowSums(sig) > 0),res.pvals,res.sig)
res.sig.wmotif <- lapply(res.pvals.wmotif, function(x) x <0.05)
# pvals.sig.filter <- pvals.dn.sig[rowSums(pvals.dn.sig) > 0,]
```

How many downregulated genes have at least one T-box transcription factors binding motifs?
```{r}
res.n.wmotif <- lapply(res.pvals.wmotif, function(x) dim(x)[1])
print(res.n.wmotif)
#dim(pvals.sig.filter)[1] # 102
```
```
$`res.dn`
[1] 102

$res.up
[1] 88

$res.reg
[1] 190

```

What fraction of the downregulated genes have at least one T-box motif?
```{r}
lapply(res.sig, function(x) round(mean(rowSums(x) > 0) * 100 , 2))
#round(dim(pvals.sig.filter)[1] / dim(pvals.dn.mat)[1] * 100,2) # 25.37%
```
```
$`res.dn`
[1] 25.37

$res.up
[1] 20.09

$res.reg
[1] 22.62

```
What is the distribution of the T-box motifs in the downregulated genes?
```{r}
lapply(res.sig.wmotif, rowSums) %>% 
  map(~stack(setNames(.x,1:length(.x)))) %>% 
  bind_rows(.id="name") %>% 
  as.tibble() %>%
  ggplot(aes(x = values)) + geom_histogram(binwidth = 1) + facet_grid(~name)

# tibble(gnames = rownames(pvals.sig.filter),
#        n_motifs = rowSums(pvals.sig.filter)) %>%
#   group_by(n_motifs) %>%
#   summarise(n_genes = n()) %>%
#   ggplot(aes(x = n_motifs, y = n_genes)) + geom_col(position = "identity")
```

## Correlations of *p*-values in *anc-1* regulated genes with at least one T-box motif
Is there a correlation in motif presence between any two motifs? Removing all the genes that don't possess a T-box transcription factors motif.
```{r}
res.cor.wmotif <- lapply(res.pvals.wmotif, Hmisc::rcorr)
print(res.cor.wmotif)

# pvals.mat.filt <- pvals.dn.mat[rowSums(pvals.dn.sig) > 0,]
# pvals.cor.filt <- Hmisc::rcorr(pvals.mat.filt)
```
```
$`res.dn`
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.26   0.12   0.12   0.27
TBX-38   0.26   1.00   0.02   0.08   0.83
TBX-39   0.12   0.02   1.00   0.63   0.03
TBX-40   0.12   0.08   0.63   1.00   0.08
TBX-43   0.27   0.83   0.03   0.08   1.00

n= 402 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.0154 0.0133 0.0000
TBX-38 0.0000        0.6263 0.1289 0.0000
TBX-39 0.0154 0.6263        0.0000 0.4903
TBX-40 0.0133 0.1289 0.0000        0.1020
TBX-43 0.0000 0.0000 0.4903 0.1020       

$res.up
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.22   0.02   0.00   0.22
TBX-38   0.22   1.00   0.02   0.00   0.82
TBX-39   0.02   0.02   1.00   0.51   0.03
TBX-40   0.00   0.00   0.51   1.00   0.07
TBX-43   0.22   0.82   0.03   0.07   1.00

n= 438 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.7065 0.9966 0.0000
TBX-38 0.0000        0.6427 0.9335 0.0000
TBX-39 0.7065 0.6427        0.0000 0.4724
TBX-40 0.9966 0.9335 0.0000        0.1302
TBX-43 0.0000 0.0000 0.4724 0.1302       

$res.reg
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33   1.00   0.25   0.07   0.06   0.25
TBX-38   0.25   1.00   0.03   0.05   0.82
TBX-39   0.07   0.03   1.00   0.57   0.04
TBX-40   0.06   0.05   0.57   1.00   0.08
TBX-43   0.25   0.82   0.04   0.08   1.00

n= 840 


P
       TBX-33 TBX-38 TBX-39 TBX-40 TBX-43
TBX-33        0.0000 0.0391 0.0604 0.0000
TBX-38 0.0000        0.4365 0.1826 0.0000
TBX-39 0.0391 0.4365        0.0000 0.2963
TBX-40 0.0604 0.1826 0.0000        0.0208
TBX-43 0.0000 0.0000 0.2963 0.0208       

```
TBX-38 and TBX-43 correlate. TBX-39 and TBX-40 correlate. Both as expected from their motifs. TBX-33 weakly correlates with all the other four transcription factors.

Plot correlations of p-values matrix.
```{r}
lapply(names(res.cor.wmotif), function(x) {
  pdf(file = paste0("experiments/2018-10-28-tbx-enrich-corr/results/tbx-motifs-cor-w-motif-",x,".pdf"),
      height = 2, width = 2, pointsize = 7)
  f_corrplot(res.cor.wmotif[[x]])
  dev.off()
})
```

Plot correlations matrices, scatter plots and statistical significance of correlations.
```{r}
f_cormatplot <- function(x) {
  pairs.panels(-log10(x),
               scale = FALSE,
               density = FALSE,
               ellipses = FALSE,
               hist.col = "grey",
               rug = FALSE,
               breaks = 10,
               stars = TRUE,
               lm = FALSE)
}

lapply(names(res.pvals.wmotif), function(x) {
  pdf(file = paste0("experiments/2018-10-28-tbx-enrich-corr/results/tbx-motifs-cormat-w-motif-",x,".pdf"),
      height = 3, width = 3, pointsize = 7)
  f_cormatplot(res.pvals.wmotif[[x]])
  dev.off()
})
```
