---
title: "Enrichment of the binding motifs of T-box transcription factors or any transcription factors in the promoters of SCF complex genes"
author: "Amir Levine"
date: "2018-12-18"
output: md_document
---

# Loading packages and data pre-processing
```{r}
if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(PWMEnrich)) {install.packages("PWMEnrich"); library(PWMEnrich)}

# Working directory
path = "experiments/2018-12-18-tbx-motifs-enrich-SCF/"

# All C. elegans promoters
load("data/mRNA-seqs-promoters.rdata")
```

Obtain SCF complex genes IDs and promoter sequences
```{r}
scf <- "skr-|rbx-1|cul-1|cul-6|sel-10|lin-23|skpt-1"
scf_gene_idx <- mRNA.ids$ensembl_gene_id[grep(scf,mRNA.ids$external_gene_name)]

scf_prs <- prs[names(prs) %in% scf_gene_idx]
```

# Enrichment of T-box transcription factors binding motifs
Load the background list of the T-box transcription factors binding motifs.
```{r}
load(paste0(path,"data/backgound-celegans-proteincoding-tbx.rdata"))
```

Motif enrichment analysis of SCF complex genes
```{r}
useBigMemoryPWMEnrich(TRUE) # using large amount of memory
res_scf = motifEnrichment(scf_prs, bg.custom)
useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
```

Generate table of TFs that are enriched in the promoters of SCF complex genes
```{r}
report_scf <- groupReport(res_scf,by.top.motifs=FALSE)
plot(report_scf)
```

Export the table of scores.
```{r}
pdf(file = paste0(path,"results/TBX-motifs-enrich-SCF.pdf"),
    height = 1.5,
    width = 5)
plot(report_scf, fontsize=7, id.fontsize=6)
dev.off()
```

Save the enrichment file
```{r}
save("scf_prs","report_scf",
     file = paste0(path,"data/TBX-motifs-enrich-SCF.Rdata"))
```

# Enrichment of all the available transcription factors binding motifs
Load the background list of the T-box transcription factors binding motifs.
```{r}
load(paste0(path,"data/background-celegans-TF.rdata"))
```

Motif enrichment analysis of SCF complex genes
```{r}
useBigMemoryPWMEnrich(TRUE) # using large amount of memory
res_scf = motifEnrichment(scf_prs, bg.custom)
useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
```

Generate table of TFs that are enriched in the promoters of SCF complex genes
```{r}
report_scf <- groupReport(res_scf,by.top.motifs=FALSE)
```

Were  the enriched transcription factors differntially regulated in anc-1 treated animals
```{r}
# The enriched transcription factors which are in the differentially expressed gene list of anc-1 regulated genes
tf_sig_a1_idx <- BOTH$Gene_name %in% report_scf[report_scf$p.value < 0.1]$id
# What were their expression levels in anc-1 treated animals
BOTH[tf_sig_a1_idx,]
# Show only the ones that had a padj < 0.1
BOTH[tf_sig_a1_idx & BOTH$padj<.1,]
```

Export the table of scores. p-value < 0.1
```{r}
pdf(file = paste0(path,"results/TF-all-motifs-enrich-SCF.pdf"),
    height = 4,
    width = 5)
plot(report_scf[report_scf$p.value < 0.05], fontsize=7, id.fontsize=6)
dev.off()
```

Save the enrichment file
```{r}
save("scf_prs","report_scf",
     file = paste0(path,"data/TF-all-motifs-enrich-SCF.Rdata"))
```