---
title: "Choosing candidates for qPCR validation of T-box targets"
author: "Amir Levine"
date: "2019-01-09"
output: md_document
---

## Load packages and datasets.
```{r}
# Data pre-processing
if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}

load("data/geneExpression.rdata")
# PWMEnrich object of Tbox transcrption factors motifs enrichement in the promoters of anc-1 RNAi regulateed genes.
load("experiments/2018-07-11-tbx-motifs/data/Tbox-anc1-reg-enriched.Rdata")
```

## Data Pre-Processing
Focus on the genes that are down-regulated by *anc-1*, as those are enriched for TBX motifs.
```{r}
# Separate AM140 and N2 data. Data that may be relevant for clustering.
AM <- BOTH %>% select(Gene_stable_ID,Gene_name,strain,log2FoldChange,padj) %>%
  filter(strain == "AM140")
NT <- BOTH %>% select(Gene_stable_ID,Gene_name,strain,log2FoldChange,padj) %>% 
  filter(strain == "N2")
AM.NT <- full_join(NT,AM,by = c('Gene_stable_ID','Gene_name'), suffix = c(".N2",".AM140"))

# Create p-values matrix. Obtain the *p*-values of motifs per sequence. Rows correspond to the different input sequences and the columns correspond to motifs. `sequence.nobg` contains the raw affinity scores.`sequence.bg` contains the corresponding P-values.
tbx.pv.dn.mat <- res.dn$sequence.bg

# Set the gene names to serve as the background list for later annotation analysis
bg <- rownames(tbx.pv.dn.mat)

# Rename the column names to be the T-box protein names.
colnames(tbx.pv.dn.mat) <- c("TBX-33","TBX-38","TBX-39","TBX-40","TBX-43")

# Covert the p-values matrix to a tibble to merge with the previous data-object.
tbx.pv.dn.tbl <- as.tibble(tbx.pv.dn.mat) %>%
  mutate(Gene_stable_ID = rownames((tbx.pv.dn.mat)))

# Combine the TBX binding-sties p-values with the previous data-object.
AM.NT <- inner_join(AM.NT, tbx.pv.dn.tbl, by='Gene_stable_ID')
```

## Regulation by TBX-38
```{r}
AM.NT %>% filter(grepl("^fb|^skr|^tbx",Gene_name)) %>%
  select(Gene_name,
         log2FoldChange.N2,
         padj.N2,
         log2FoldChange.AM140,
         padj.AM140,
         starts_with("TBX")) %>%
  ggplot(aes(x = log2FoldChange.N2,
             y = -log10(`TBX-38`),
             label = Gene_name)) + 
  geom_point() + 
  geom_text() +
  geom_hline(aes(yintercept= -log10(0.05) )) + # p-value threshold < 0.05
  geom_vline(aes(xintercept=-1))  # fold-change threshold.
```
Genes that are regulated by TBX-38 (p<0.1) with fold-change > 1:
- tbx-37
- tbx-43

- fbxc-25

## Regulation by TBX-39
```{r}
AM.NT %>% filter(grepl("^fb|^skr|^tbx",Gene_name)) %>%
  select(Gene_name,
         log2FoldChange.N2,
         padj.N2,
         log2FoldChange.AM140,
         padj.AM140,
         starts_with("TBX")) %>%
  ggplot(aes(x = log2FoldChange.N2,
             y = -log10(`TBX-39`),
             label = Gene_name)) + 
  geom_point() + 
  geom_text() +
  geom_hline(aes(yintercept=  -log10(0.05) )) + # p-value threshold.
  geom_vline(aes(xintercept=-1))  # fold-change threshold.
```
Genes that are regulated by TBX-39 (p<0.1) with fold-change > 1:
- tbx-37
- tbx-38

## Regulation by TBX-40
```{r}
AM.NT %>% filter(grepl("^fb|^skr|^tbx",Gene_name)) %>%
  select(Gene_name,
         log2FoldChange.N2,
         padj.N2,
         log2FoldChange.AM140,
         padj.AM140,
         starts_with("TBX")) %>%
  ggplot(aes(x = log2FoldChange.N2,
             y = -log10(`TBX-40`),
             label = Gene_name)) + 
  geom_point() + 
  geom_text() +
  geom_hline(aes(yintercept= -log10(0.05)  )) + # p-value threshold.
  geom_vline(aes(xintercept=-1))  # fold-change threshold.
```
Genes that are regulated by TBX-40 (p<0.1) with fold-change > 1:
- tbx-37
- tbx-38
- tbx-11

- skr-8
- skr-15

## Genes to check their expression levels when knocking down various tbx genes
- tbx-37
- tbx-38
- tbx-43
- tbx-11
(must choose 2, max 3 tbx genes)

- skr-8
- skr-15

- fbxc-25
- fbxc-56


Focusing on highly significant genes first-
- tbx-38 (make sure no overlap)
- tbx-43
- skr-8
- fbxc-25
