---
title: "Check for an enrichment of C. elegans transcription factors motifs in the promoter of anc-1 targets"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Load bioconductor and packages
```{r}
source("https://bioconductor.org/biocLite.R")
if (!require(tidyverse)) {install.packages("tidyverse"); library(tidyverse)}
if (!require(PWMEnrich)) {install.packages("PWMEnrich"); library(PWMEnrich)}
```

# Gene-lists and Sequences
Load data from upstream analyses.
```{r}
# Load CEL-seq gene lists
load("data/geneExpression.rdata")
# Load promoters of C. elegans genes. Data generated by "Celegans-promoters.R"
load("data/Celegans-promoters.rdata")
# Load the background PWMs applied to all sequences object, which is generated by "TFs-background.Rmd".
load("experiments/2018-09-02-TFs-motifs/data/background-celegans-TF.rdata")
```


# Motif enrichment scanning
Perform transcription factors enrichment analysis to sequences regulated by *anc-1*, in down-regulated and up-regulated genes of each strain.
Sequences of interest (goi) are the promoter sequences of all the genes that are significantly (p<0.1) regulated by *anc-1*, in both strains.
Set all genes lists of interest.
Define function for motif enrichment analysis based on the bioconductor (PWMEnrich  package)[https://www.bioconductor.org/packages/release/bioc/html/PWMEnrich.html]
Apply function to each gene list and store results as a list file.
Long run, save the generated object.
```{r}
reg.genes <- list(N2.up = DEG$N2.up,
                  N2.dn = DEG$N2.dn,
                  AM.up = DEG$AM.up,
                  AM.dn = DEG$AM.dn)

me <- function(goi,mRNA.seqs.promoter,prs) {
  mRNA.seqs.promoter$CELseq.hits <- mRNA.seqs.promoter$ensembl_gene_id %in% goi
  return(motifEnrichment(prs[mRNA.seqs.promoter$CELseq.hits], bg.custom))
}

useBigMemoryPWMEnrich(TRUE) # using large amount of memory
#registerCoresPWMEnrich(4) # Parallel execution Not supported in Windows

res <- lapply(reg.genes, me,mRNA.seqs.promoter,prs)

useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
#registerCoresPWMEnrich(NULL) # disable Parallel execution

save(res,file = "experiments/2018-09-02-TFs-motifs/data/TFs-motif-enrichment-results")
```

Previous analysis focused on the intersect of the two strains, with upregulated and downregulated genes analyzed separately. 
```{r}
goi <- DEG$INTERSECT.dn # CEL-seq DEGs, query genes
length(goi) # 93 downregulated genes in both strains.
mRNA.seqs.promoter <- mRNA.seqs.promoter %>% 
  mutate(CELseq.hits = ensembl_gene_id %in% goi)
```

## Random sequences as a control
```{r}
set.seed(168)
goi <- sample(mRNA.seqs.promoter$ensembl_gene_id,168)
mRNA.seqs.promoter <- mRNA.seqs.promoter %>% 
  mutate(CELseq.hits = ensembl_gene_id %in% goi)
useBigMemoryPWMEnrich(TRUE) # using large amount of memory
res = motifEnrichment(prs[mRNA.seqs.promoter$CELseq.hits], bg.custom)
useBigMemoryPWMEnrich(FALSE) # disable using large amount of memory
report <- groupReport(res,by.top.motifs=TRUE) # Report of all enrichment scores.
plot(report[report$p.value < 0.05], fontsize=7, id.fontsize=6)
```


